Inference methods comparison for the simplified Naomi model using the Kolmogorov-Smirnov test statistic

### 2023-05-04 src/naomi-simple_ks/ks.Rmd
Use random_only TMB and reorganise KS comparison notebook #27

---
title: "Inference methods comparison for the simplified Naomi model"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background**  We have run the simplified Naomi model using a range of inference methods: `TMB`, `aghq`, `adam` and `tmbstan`.

@@

We create histograms and empirical cumulative distribution function (ECDF) difference plots of the samples from each method.
All of the possible parameter names are as follows:

```{r}
pars <- names(tmb$fit$sample)
pars
```

We will produce plots about the following subset of them.
There is no particular reason to choose this subset rather than other, it's quite arbitrary.

```{r}
pars_eval <- pars %in% c("beta_rho", "beta_alpha", "beta_lambda", "beta_anc_rho", "beta_anc_alpha", "u_rho_x", "us_rho_x", "u_rho_xs")
names(pars_eval) <- pars
```

The `beta_rho` and `beta_anc_rho` parameters are used in other reports, so create dataframes here to save as artefacts:

```{r}
beta_rho <- histogram_and_ecdf("beta_rho", i = 1, return_df = TRUE)
saveRDS(beta_rho$df, "beta_rho.rds")

beta_anc_rho <- histogram_and_ecdf("beta_anc_rho", return_df = TRUE)
saveRDS(beta_anc_rho$df, "beta_anc_rho.rds")
```

@@

## Outcome variables

These are all of the outcome variables which we might be interested in to monitor:

```{r}
names(tmb$fit$sample)[!(names(tmb$fit$sample) %in% unique(names(tmb$fit$obj$env$par)))]
```
The most important ones (this has been checked with Jeff) are HIV prevalence (`rho_t1_out`), ART coverage (`alpha_t1_out`), and HIV incidence (`lambda_t1_out`).

* HIV prevalence has `r dim(tmb$fit$sample$rho_t1_out)[1]` variables (rows)
* ART coverage has `r dim(tmb$fit$sample$alpha_t1_out)[1]` variables (rows)
* HIV incidence has `r dim(tmb$fit$sample$lambda_t1_out)[1]` variables (rows)

```{r}
dim(tmb$fit$sample$rho_t1_out)
dim(tmb$fit$sample$alpha_t1_out)
dim(tmb$fit$sample$lambda_t1_out)
```

Within `tmb$naomi_data` there is a dataframe called `mf_out` (modelframe output) which contains the area, age, sex mapping these `r dim(tmb$fit$sample$rho_t1_out)[1]` rows.
As a sanity check, confirm that the number of areas times number of ages times number of sexes indeed equals the number of rows:

It makes sense only to assess inferential accuracy at the finest level so as to avoid double counting.

```{r}
mf_out$area_id %>% unique()
mf_out$sex %>% unique()
mf_out$age_group %>% unique()

mf_out_fine <- mf_out %>%
  tibble::rownames_to_column("id") %>%
  mutate(id = as.numeric(id)) %>%
  filter(
    area_id %in% paste0("MWI_4_", 1:32, "_demo"),
    sex %in% c("male", "female"),
    age_group %in%
      c(
        "Y000_004", "Y005_009", "Y010_014", "Y015_019", "Y020_024", "Y025_029",
        "Y025_034", "Y030_034", "Y035_039", "Y040_044", "Y045_049", "Y050_054",
        "Y055_059", "Y060_064", "Y065_069", "Y070_074", "Y075_079", "Y080_999"
      )
  )
```

An `id` column has been added to `mf_out_fine` to enable merging to the samples.

```{r}
head(mf_out_fine)
nrow(mf_out_fine)
```

```{r}
rho_t1_out_fine <- tmb$fit$sample$rho_t1_out[mf_out_fine$id, ]
dim(rho_t1_out_fine)
rho_t1_out_fine[1:10, 1:10]
```

Something weird happening with `adam`, go back and check this:

```{r}
histogram_and_ecdf(par = "rho_t1_out", i = 1)
```

@@

```{r}
ks_helper <- function(par, starts_with = FALSE, ...) {
  to_ks_df(par, starts_with = starts_with) %>% ks_plot(par, ...)
}
```

@@

### Outcome variables

```{r}
ks_df_out <- function(par) {
  to_ks_df(par = par, outputs = TRUE, id = mf_out_fine$id) %>%
  select(-par) %>%
  left_join(
    mf_out_fine %>%
      rename("full_id" = "id") %>%
      tibble::rowid_to_column("index") %>%
      mutate(index = as.numeric(index))
  )
}

ks_rho_t1_out <- ks_df_out(par = "rho_t1_out")
ks_alpha_t1_out <- ks_df_out(par = "alpha_t1_out")
ks_lambda_t1_out <- ks_df_out(par = "lambda_t1_out")

ks_plot(ks_df = ks_rho_t1_out, par = "rho_t1_out", alpha = 0.2)
ks_plot(ks_df = ks_alpha_t1_out, par = "alpha_t1_out", alpha = 0.2)
ks_plot(ks_df = ks_lambda_t1_out, par = "lambda_t1_out", alpha = 0.2)
```

What is going on with those very low values of KS?
It's that there are new new infections in those age groups, and the posterior samples from each method are exactly the same, in particular draws of only zero.

@@

## Summary table

@@

ks_summary %>%
  gt::gt() %>%
  gt::fmt_number(
    columns = starts_with("KS"),
    decimals = 3
  ) %>%
  gt::tab_style(
    style = cell_fill(color = "#0072B2"),
    locations = cells_body(
      columns = `KS(adam, tmbstan)`,
      rows = `KS(adam, tmbstan)` < `KS(aghq, tmbstan)` & `KS(adam, tmbstan)` < `KS(TMB, tmbstan)`
    )
  ) %>%
    gt::tab_style(
    style = cell_fill(color = "#0072B2"),
    locations = cells_body(
      columns = `KS(aghq, tmbstan)`,
      rows = `KS(aghq, tmbstan)` < `KS(adam, tmbstan)` & `KS(aghq, tmbstan)` < `KS(TMB, tmbstan)`
    )
  ) %>%
    gt::tab_style(
    style = cell_fill(color = "#0072B2"),
    locations = cells_body(
      columns = `KS(TMB, tmbstan)`,
      rows = `KS(TMB, tmbstan)` < `KS(aghq, tmbstan)` & `KS(TMB, tmbstan)` < `KS(adam, tmbstan)`
    )
  )
```

```{r}
ks_summary %>%
  filter(Type != "Hyper") %>%
  ks_plot_many("TMB", "aghq")

(ks_summary_out <- ks_summary %>%
  group_by(Type) %>%
  summarise(
    TMB = mean(`KS(TMB, tmbstan)`),
    aghq = mean(`KS(aghq, tmbstan)`),
    adam = mean(`KS(adam, tmbstan)`)
  )) %>%
  gt::gt()
```

Save some things for output:

```{r}
saveRDS(ks_summary_out, "ks_summary_out.rds")
```

@@

Nodes where `TMB` beats `aghq`:

```{r}
ks_comparison(ks_summary, method1 = "aghq", method2 = "TMB", 10)
```

```{r}
histogram_and_ecdf(par = "ui_anc_alpha_x", i = 18)
histogram_and_ecdf(par = "u_alpha_xs", i = 18)
histogram_and_ecdf(par = "ui_anc_alpha_x", i = 29)
```

Nodes where `aghq` beats `TMB`:

```{r}
ks_comparison(ks_summary, method1 = "TMB", method2 = "aghq", 10)  
```

```{r}
histogram_and_ecdf(par = "us_alpha_xs", i = 30)
histogram_and_ecdf(par = "us_alpha_xs", i = 31)
histogram_and_ecdf(par = "us_alpha_xs", i = 26)
```

>>>>

---
title: "Inference methods comparison for the simplified Naomi model using the Kolmogorov-Smirnov test statistic"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background**  We have run the simplified Naomi model using a range of inference methods: `TMB`, `aghq`, `adam` and `tmbstan`.

@@

We create histograms and empirical cumulative distribution function (ECDF) difference plots of the samples from each method.
All of the possible latent field and hyperparamter names are given by `pars`:

```{r}
pars <- names(tmb$fit$sample)
```

There are also all of the Naomi outcome variables:

```{r}
names(tmb$fit$sample)[!(names(tmb$fit$sample) %in% unique(names(tmb$fit$obj$env$par)))]
```

We will especially focus on the outputs:

* HIV prevalence (`rho_t1_out`) which has `r dim(tmb$fit$sample$rho_t1_out)[1]` variables (rows)
* ART coverage (`alpha_t1_out`) which has `r dim(tmb$fit$sample$alpha_t1_out)[1]` variables (rows)
* HIV incidence (`lambda_t1_out`) which has `r dim(tmb$fit$sample$lambda_t1_out)[1]` variables (rows)

These outputs are at an aggregate resolution as well as at the finest resolution.
It seems to make sense to only to assess inferential accuracy at the finest resolution so as to avoid double counting.
As such, we need to find the subset of these variables at the finest resolution.
Within `tmb$naomi_data` there is a dataframe called `mf_out` (model frame output) which contains the area, age, sex mapping for these `r dim(tmb$fit$sample$rho_t1_out)[1]` rows.
As a sanity check, confirm that the number of areas times number of ages times number of sexes indeed equals the number of rows:

The unique levels for area, age and sex are:

```{r}
mf_out$area_id %>% unique()
mf_out$age_group %>% unique()
mf_out$sex %>% unique()
```

Filter to only the finest resolution, adding an `id` column has to enable merging to the samples:

```{r}
mf_out_fine <- mf_out %>%
  tibble::rownames_to_column("id") %>%
  mutate(id = as.numeric(id)) %>%
  filter(
    area_id %in% paste0("MWI_4_", 1:32, "_demo"),
    sex %in% c("male", "female"),
    age_group %in%
      c(
        "Y000_004", "Y005_009", "Y010_014", "Y015_019", "Y020_024", "Y025_029",
        "Y025_034", "Y030_034", "Y035_039", "Y040_044", "Y045_049", "Y050_054",
        "Y055_059", "Y060_064", "Y065_069", "Y070_074", "Y075_079", "Y080_999"
      )
  )
```

We will produce plots about the a small subset of parameters for the time being.
There is no particular reason to choose this subset rather than other, it's quite arbitrary.

```{r}
pars_eval <- pars %in% c("beta_rho", "u_rho_x")
names(pars_eval) <- pars
```

@@

## `rho_t1_out`

There are too many variables to plot this here!

## `alpha_t1_out`

There are too many variables to plot this here!

## `lambda_t1_out`

There are too many variables to plot this here!

@@

```{r}
ks_helper <- function(par, starts_with = FALSE, ...) {
  to_ks_df(par, starts_with = starts_with) %>% ks_plot(par, ...)
}

ks_df_out <- function(par) {
  to_ks_df(par = par, outputs = TRUE, id = mf_out_fine$id) %>%
  select(-par) %>%
  left_join(
    mf_out_fine %>%
      rename("full_id" = "id") %>%
      tibble::rowid_to_column("index") %>%
      mutate(index = as.numeric(index))
  )
}

```

@@

### `rho_t1_out`

```{r}
ks_df_out(par = "rho_t1_out") %>%
  ks_plot(par = "rho_t1_out", alpha = 0.2)
```

### `alpha_t1_out`

```{r}
ks_df_out(par = "alpha_t1_out") %>%
  ks_plot(par = "rho_t1_out", alpha = 0.2)
```

### `lambda_t1_out`

```{r}
ks_df_out(par = "rho_t1_out") %>%
  filter(!age_group %in% c("Y005_009", "Y010_014", "Y080_999")) %>%
  ks_plot(par = "lambda_t1_out", alpha = 0.2)
```

Filter out the age group 5-9, 10-14 and 80+ here becausethere are new new infections in those age groups, and the posterior samples from each method are exactly the same: in particular draws of identically zero.

@@

## Summary {.tabset .tabset-fade}

@@

saveRDS(ks_summary_out, "ks_summary_out.rds")

ks_summary %>%
  filter(Type != "Hyper") %>%
  ks_plot_many("TMB", "aghq")
```

```{r}
(ks_summary_out <- ks_summary %>%
  group_by(Type) %>%
  summarise(
    TMB = mean(`KS(TMB, tmbstan)`),
    aghq = mean(`KS(aghq, tmbstan)`),
    adam = mean(`KS(adam, tmbstan)`)
  )) %>%
  gt::gt()
```

### Latent field

```{r}
ks_table(ks_summary, type = "Latent field")
```

### Hyperparameters

```{r}
ks_table(ks_summary, type = "Latent field")
```

### Outputs

```{r}
#' To-do
```

@@

### Nodes where `TMB` beats `aghq`

```{r}
ks_comparison(ks_summary, method1 = "aghq", method2 = "TMB", 10)

histogram_and_ecdf(par = "us_rho_x", i = 19)
histogram_and_ecdf(par = "ui_anc_rho_x", i = 2)
histogram_and_ecdf(par = "u_rho_x", i = 2)
```

### Nodes where `aghq` beats `TMB`

```{r}
ks_comparison(ks_summary, method1 = "TMB", method2 = "aghq", 10)

histogram_and_ecdf(par = "us_alpha_xs", i = 18)
histogram_and_ecdf(par = "us_alpha_xs", i = 30)
histogram_and_ecdf(par = "u_alpha_xa", i = 10)
```
